---
title: "Introduction to R, part 3d"
author: "Steve Simon"
date: "Saturday, August 13, 2016"
output: html_document
---

This program is part of a series of R programs for the Introduction to R class. This program shows you how to calculated group means and standard deviations.

```{r r-version-and-current-date}
R.version.string
Sys.Date()
```

Erase any stray data and load the data from part 3c.

```{r read-part-3c-data}
rm(list=ls())
load(file="part3c.RData")
```

You should also calculate the means for each continuous variable within each level of the categorical variable.

Note: you don't really need the na.rm=TRUE argument here, but you might in other examples.

```{r pu.by.fev}
# fev mean by smoke.factor
by(pu$fev,pu$smoke.factor,mean,na.rm=TRUE)
# fev standard deviation by smoke.factor
by(pu$fev,pu$smoke.factor,sd,na.rm=TRUE)
```

The output is not labelled all that clearly, but you can use the paste function to make things look nicer.

```{r pu.paste.fev}
grp.means <- by(pu$fev,pu$smoke.factor,mean)
grp.stdev <- by(pu$fev,pu$smoke.factor,sd)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means),colon,round(grp.means,1),
      plus.minus,round(grp.stdev,1),sep="")
```

Let's calculate mean ages for smokers and nonsmokers. Notice that smokers are 4 years older on average than nonsmokers.

```{r pu.paste.age}
grp.means <- by(pu$age,pu$smoke.factor,mean)
grp.stdev <- by(pu$age,pu$smoke.factor,sd)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means),colon,round(grp.means,1),
      plus.minus,round(grp.stdev,1),sep="")
```

## On your own

Compute the mean and standard deviation for fev by sex.

## Read in the housing data set

You should find the data file on housing prices in Albuquerque at

--> http://lib.stat.cmu.edu/DASL/Datafiles/homedat.html

You should cut-and-paste the actual data to notepad and save it with the name housing.txt. You need to read in the file using the read.table function.

For technical reasons, you want the as.is=TRUE argument. Otherwise AGE and TAX are converted to factors which causes all sorts of grief.

```{r read.housing}
al <- read.table(file="housing.txt",header=TRUE,as.is=TRUE)
head(al)
tail(al)
```

```{r univariate}
summary(al)
```

There is a problem with AGE. The person who entered the data intended for an asterisk to represent a missing value. But R interpreted the asterisk as a string and converted all the numeric ages to strings also. That's why R summarizes age using counts rather than means. Use the as.numeric function to fix this.

A side effect, which is desired in this case, is that anything that cannot be converted from a string to a number (like an asterisk) is instead converted to a missing value.

```{r convert.age}
al$age.years <- as.numeric(al$AGE)
summary(al$age.years)
```

There is a mix of categorical and continuous variables. Recall that you use a scatterplot to examine the relationship between two continuous variables, a crosstabls and an odds ratio to examine the relationship between two categorical variables, and a boxplot to examine the relationship between a categorical variable and a continuous variable.

## On your own

1. Convert the asterisks in the TAX variable to missing. How many missing values are there for AGE and for TAX?

2. Create factors for NE, CUST, and COR. Draw bar charts for each of these factors.

3. Find the largest house (biggest SQFT) in the data set. Is the largest house also the most expensive house?

4. Calculate frequency counts for FEATS. Are there any houses with the no features? Are there any houses with every possible feature?

5. Evaluate the relationship between PRICE and SQFT using a scatterplot. Include a smooth curve. Do larger houses tend to cost more?

6. Evaluate the relationship between CUST and PRICE using a boxplot. Calculate the mean price and standard deviation by CUST.

7. What is the difference in average prices between northeast houses and other houses? What is the difference in average taxes?

8. Are custombuilt houses more likely to appear on corner lots? Calculate the percentages and compute a relative risk.

9. Place all the key results into a Word document, add a bit of commentary, and turn it in.

## A third data set

If there is time, we can explore a third data set, diamond prices. The data is at 

--> http://www.amstat.org/publications/jse/v9n2/4cdata.txt

and a description is at 

--> http://www.amstat.org/publications/jse/v9n2/4c.txt

## Save all my work

```{r save-image}
save.image("part3d.RData")
```