---
title: "Introduction to R, part 3a"
author: "Steve Simon"
date: "Saturday, August 13, 2016"
output: html_document
---

# Part 3a. Read in data and assign names

This program illustrates how to import and manipulate simple data sets in R and how to produce some simple descriptive statistics. Included in this program are some exercises for you to do on your own.

Start with a clean data set.

```{r clean-slate}
rm(list=ls())
```

Today, you will analyze some data sets that have a mix of categorical and continuous variables.

## read in the pu data set

```{r read.pu}
fn <- "http://www.amstat.org/publications/jse/datasets/fev.dat.txt"
pu <- read.table(file=fn)
head(pu)
tail(pu)
```

Since the file looks good, let's save a local copy.

```{r write-pu}
write.csv(pu,file="fev.txt",row.names=FALSE,col.names=FALSE)
```

## Assign names and create factors

You should assign names to each column of data and convert the columns with numeric codes into factors

```{r factors.pu}
names(pu) <- c("age","fev","ht","sex","smoke")
pu$smoke.factor <- factor(pu$smoke,levels=0:1,
                          labels=c("nonsmoker","smoker"))
```

## On your own

Create a factor for sex.

## Save all my work

```{r save-image}
save.image("part3a.RData")
```
