---
title: "Introduction to R, part 3b"
author: "Steve Simon"
date: "Saturday, August 13, 2016"
output: html_document
---

This program is part of a series of R programs for the Introduction to R class. This program reviews scatterplots, correlations, and crosstabulations.

```{r r-version-and-current-date}
R.version.string
Sys.Date()
```

Erase any stray data and load the data from part 3a.

```{r read-part-3a-data}
rm(list=ls())
load(file="part3a.RData")
```

## Summary for the pu variables

There is a mix of categorical and continuous variables in this data set, but the summary function should work for all of them.

```{r summary.pu}
summary(pu)
```

## On your own

Calculate correlations among age, ht, and fev.

## Scatterplot

Recall that you use a scatterplot to examine the relationship between two continuous variables.

```{r pu.scatterplot}
plot(pu$ht,pu$fev)
smooth.curve <- lowess(pu$ht,pu$fev)
lines(smooth.curve)
```

## On your own

Draw a scatterplot for age and fev. Add a linear regression trend line to this graph.

## crosstabs

Also recall that you use a crosstabulation to examine the relationship between two categorical variables.

```{r pu.crosstabs}
table(pu$sex.factor,pu$smoke.factor)
prop.table(table(pu$sex,pu$smoke.factor),margin=1)
round(100*prop.table(table(pu$sex.factor,pu$smoke.factor),margin=1))
```

## Save all my work

```{r save-image}
save.image("part3b.RData")
```
