---
title: "Introduction to R, part 3a"
author: "Steve Simon"
date: "Saturday, August 13, 2016"
output: html_document
---

This program is part of a series of R programs for the Introduction to R class. This program shows you how to read the text file that you need for most of Part 3.

```{r r-version-and-current-date}
R.version.string
Sys.Date()
```

Start with a clean data set.

```{r clean-slate}
rm(list=ls())
```

Today, you will analyze some data sets that have a mix of categorical and continuous variables. The first data set looks at pumonary function in a group of children. You can find a description of this data set at

--> http://ww2.amstat.org/publications/jse/datasets/fev.txt

## read in the pu data set

```{r read.pu}
fn <- "http://www.amstat.org/publications/jse/datasets/fev.dat.txt"
pu <- read.table(file=fn)
head(pu)
tail(pu)
```

Since the file looks good, let's save a local copy.

```{r write-pu}
write.csv(pu,file="fev.txt",row.names=FALSE,col.names=FALSE)
```

## Assign names and create factors

You should assign names to each column of data and convert the columns with numeric codes into factors

```{r factors.pu}
names(pu) <- c("age","fev","ht","sex","smoke")
pu$smoke.factor <- factor(pu$smoke, levels=0:1,    labels=c("nonsmoker", "smoker"))
```

## On your own

Create a factor for sex.

## Don't peek below until you've done this yourself.

```{r additional-factors}
pu$sex.factor <- factor(pu$sex, levels=0:1,    labels=c("female", "male"))
```

## Save all my work

```{r save-image}
save.image("part3a.RData")
```
